MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-23 chapters/MetaTracts.tex
Updated text ready for review.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Corouge2004, Brun2003, Brun2004, Zhang2008, westinMEDIA02}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{jonasson2005, ODonnell2007, Brun2004}.

@@

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are joined. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores: Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers.

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the magnified version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps. MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigen values of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).  The following measures are defined in their work~\cite{Frangi1998}.  

@@

In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure. Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors. Thus we compute a measure, ``equal" to the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.

@@

In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. Figure~\ref{fig:reliable_hessian}A shows the entire data set. Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively. Figure~\ref{fig:reliable_hessian}D shows a close up (region similar to Figure~\ref{fig:data-char}D). Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue(Fig.~\ref{fig:algo}B). Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Fig.~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=0.4\textwidth]{imagesMT2014/image4}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

\begin{itemize}
\item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
\item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible. Thus, we want the distance between $C_p$ and the start point for the next cylinder to be as large as possible. We measure distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$. For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B). We refer to this perpendicular projection distance as \textit{``projected\_dist"}. 
\end{itemize}
% \begin{equation}\label{eqn:algo_1}
% d=\frac{1}{3}e^{(\frac{-angle^{2}}{\alpha^{2}})}+\frac{2}{3}(1-e^{\frac{-{PerpDist}^{2}}{\beta^{2}}})
% \end{equation}
We further define a priority for each of the candidate vertices based on the above factors.
For each cylinder in a MetaTract, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
\begin{equation}
Priority = \gamma_1 e^{(-angle^2 / \alpha^2)} + \gamma_2e^{(-projected\_dist^2 / \beta^2)}
\label{eqn:algo_1}
\end{equation}
$\gamma_1$, $\gamma_2$ are the  weights ($\mathbb{R}_{\ge 0}$)  which decide how the priority depends on the affine combination of the two factors. For all our cases we use ($\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$). In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$). At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps. Essentially, Equation~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction. This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.

In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then removed. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A. The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles. Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their length.
MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering. We have found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).  
Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
We found that different clustering techniques performed preferably for different measures.  The two step clustering also made the problem more tractable and  the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should be considered similar. Brun et al. ~\cite{Brun2004} based their measure on a 9-D fiber descriptors. In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure}. There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts. The edge weights give a distance matrix representing the distance between each pair of points. Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix. 
%
%August 13 2014
%For all our tests the fibers were then embedded in a $\eta$ low dimensional space. The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
%  
We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
For our test case there are two major directions of the fiber bundles. So $K$ was set to 2. It is important to note that due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern. Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. As is expected there is a clear distinction based on fiber orientation. Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.

>>>>

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering ~\cite{Moberts2005, Zhang2008} and those using spectral clustering ~\cite{ Brun2004,jonasson2005, ODonnell2007}.

@@

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}. 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores: Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the magnified version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

Figure \ref{fig:flowchart} shows our pipeline. MetaTracts generations  is explained in Section.~\ref{sec:approach} and  finding plausible fiber bundles in Section~\ref{subsec:fiber-bundles}, Section~\ref{sec:vis} talks about visualization extensions.
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$).  The following measures are defined in their work~\cite{Frangi1998}.  

@@

In equation~\ref{RB}, $\mathcal{R_{B}}$ provides a measure of deviation from a blob like structure while in equation~\ref{RA}, $\mathcal{R_{A}}$ distinguishes between plate-like and line-like structure. Grey-scale variations and close proximity of the fibers in our data make the Hessians computed at each voxels susceptible to errors. Thus we compute the ``vesselness'' measure to determine which locations in the volume provide reliable local orientation.

@@

In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. Figure~\ref{fig:reliable_hessian}A shows the entire data set. Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. Figure~\ref{fig:reliable_hessian}B and C show 2D slices along the Z and X axis respectively. Figure~\ref{fig:reliable_hessian}D shows a magnified region similar to Figure~\ref{fig:data-char}D. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) Magnified region marked in green square (in C).}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
 In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue (Fig.~\ref{fig:algo}B). Fig.~\ref{fig:algo}(C) shows a single cylinder of a MetaTracts. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=0.3\textwidth]{imagesMT2014/image4}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

\begin{itemize}
\item \textit{Orientation Similarity}: We want the orientation of the start points for the consecutive cylinders to have a similar orientations ($N_p$). 
\item \textit{Large distance}: We want the MetaTracts to traverse the data using as few cylinders as possible. Thus, we want the distance between $C_p$ and the start point for the next cylinder to be as large as possible. We measure distance of a candidate vertex from $C_p$ by projecting the Euclidean distance between them onto $N_p$. For example the Euclidean distance between the blue and the orange vertices projected onto $N_P$ gives distance (Figure~\ref{fig:algo}B). We refer to this perpendicular projection distance as \textit{``projected\_dist"}. 
\end{itemize}
% \begin{equation}\label{eqn:algo_1}
% d=\frac{1}{3}e^{(\frac{-angle^{2}}{\alpha^{2}})}+\frac{2}{3}(1-e^{\frac{-{PerpDist}^{2}}{\beta^{2}}})
% \end{equation}
We define a priority for each of the candidate vertices based on the above factors.
For each cylinder in a MetaTract, we put the \textit{candidate vertices} of it in a priority queue based on Equation \ref{eqn:algo_1}.
\begin{equation}
Priority = \gamma_1 e^{(-angle^2 / \alpha^2)} + \gamma_2e^{(-projected\_dist^2 / \beta^2)}
\label{eqn:algo_1}
\end{equation}
$\gamma_1$, $\gamma_2$ are the  weights ($\mathbb{R}_{\ge 0}$)  which decide how the priority depends on the affine combination of the two factors. For all our cases we use ($\gamma_1$ = $1 / 3 $ and $\gamma_2$ = $2 / 3$). In general we suggest $\gamma_1 +\gamma_2 = 1 $ and $\gamma_1 \leq \gamma_2$). At each iteration we pick the top element in the priority queue generate the corresponding cylinder and repeat the steps. Essentially, Equation~\ref{eqn:algo_1}, selects a grid point which is the furthest from the current point and going in a similar direction. This gives it the advantage of tackling noise/errors in local orientation better than integral curves by looking at multiple choices for vertices and avoiding intra cell interpolation in an already noisy environment.

In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then removed. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A. The MetaTracts are colored with the mean orientation direction mapped to the RGB space. Fig.\ref{fig:length_distribution}A also shows a single slice of YZ plane comparable to Fig.~\ref{fig:data-char}C and Fig.~\ref{fig:reliable_hessian}C.That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles. Fig.~\ref{fig:length_distribution}(B) shows the MetaTracts of a particular bundle colored according to their individual length.
MetaTracts in a given fiber bundle do not extend the full length of the bundle. MetaTracts may have different lengths and may partially overlap.

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length:  $2\mu m$). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles. 
Carbon fibers have both orientation and geometric proximity information, both of which can be used for clustering. Experimentally we found that orientation measure was very reliable while measures of geometric proximity had problems with fibers which partially overlap (Fig.~\ref{fig:length_distribution}B).  
Instead of creating a heuristic which artificially combines the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity.
We found that different clustering techniques performed preferably for different measures.  The two step clustering also made the problem more tractable and  the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

Brun et al.~\cite{Brun2003} based their measure on the simple assumption that two traces with similar endpoints should be considered similar. Brun et al. ~\cite{Brun2004} based their measure on a 9-D fiber descriptors. In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure}. There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts. The edge weights give a distance matrix representing the distance between each pair of points. Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix. 
%
%August 13 2014
%For all our tests the fibers were then embedded in a $\eta$ low dimensional space. The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
%  
We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
For our test case there are two major directions of the fiber bundles and $K$ was set to 2. Due to the advantages offered by the dimensionality reduction, even if there are curved fibers bundles the user has to provide just the major fiber bundle directions of the weaving pattern. Figure~\ref{fig:orientation_clustering}B shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. As is expected there is a clear distinction based on fiber orientation. Figure~\ref{fig:orientation_clustering}(A,C) shows the MetaTracts colored according to clustering results.
